Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-04-28 paper/ms.tex
Add smooth ISM, fgas models

\FloatBarrier
\newpage

@@

@@

\noindent
{\bf Smooth Interstellar Medium Model}
\\
\noindent
In addition to the SMUGGLE model, we considered a simpler model of the
interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
the multiphase nature of the interstellar medium is modeled in a subgrid
manner by allowing each resolution element to have a ``cold'' and ``hot''
component, with the equation of state of the gas suitably modified. Gas is
allowed to interchange between the ``cold'' and ``hot'' components through
processes such as cooling and stellar feedback. Cold gas is allowed to undergo
star formation. The pattern speed evolution of this setup is given in Extended
Data Fig.~X, which shows that the pattern speed evolves in a similar manner to
the SMUGGLE case shown in Fig.~\ref{fig:prop}.

@@

We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
components (initial stellar disk, stellar bulge, and gas) and a mass
resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
resolution is closest to ``level 3'' in the AURIGA
simulations.\cite{2017MNRAS.467..179G} This corresponds to approximately
$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
We used a softening length of $0.02\,\textrm{kpc}$ for all components.

\noindent
{\bf Varying Gas Fraction}
\\
\noindent
To be written.

>>>>

\FloatBarrier
\clearpage

@@

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/fig-GFM.pdf}
\caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
evolution is shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE
(orange), and smooth ISM (red) cases. The smooth ISM model is an older model
for the ISM which treats its multiphase nature in a subgrid
fashion.\cite{2003MNRAS.339..289S} This fundamentally differs from the SMUGGLE
model, which explicitly resolve the hot and cold phases of the
ISM.\cite{2019MNRAS.489.4233M} The pattern speed in the smooth ISM case is
broadly similar to the evolution in the SMUGGLE case. This shows that the
stability of the pattern speed is not simply a result of our assumed model for
the ISM.}
\label{fig:GFM}
\end{figure}

We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
components (initial stellar disk, stellar bulge, and gas) and a mass
resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
resolution is closest to ``level 3'' in the AURIGA
simulations.\cite{2017MNRAS.467..179G} This corresponds to approximately
$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
We used a softening length of $0.02\,\textrm{kpc}$ for all components.
Snapshots were saved at equal intervals of $0.005$ in the time units of the
simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-fgas.pdf}
\caption{\textbf{Pattern speed evolution with varying gas fractions.} We
explored the impact of lowering the initial gas surface density of our
fiducial disk on the evolution of the pattern speed. The surface densities we
tested of $20$, $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to
initial gas fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that
initial surface densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in
bars which remain at a constant pattern speed, while an initial surface
density of $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in
roughly the same manner as the \Nbody{} case. }
\label{fig:fgas}
\end{figure}

@@

\noindent
{\bf Smooth Interstellar Medium Model}
\\
\noindent
In addition to the SMUGGLE model, we considered a simpler model of the
interstellar medium based upon ref.\cite{2003MNRAS.339..289S} In this model,
the multiphase nature of the interstellar medium is modeled in a subgrid
manner by allowing each resolution element to have a ``cold'' and ``hot''
component, with the equation of state of the gas suitably modified. Gas is
allowed to interchange between the ``cold'' and ``hot'' components through
processes such as cooling and stellar feedback. Cold gas is allowed to undergo
star formation. The pattern speed evolution of this setup is given in Extended
Data Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a
similar manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.

@@

\noindent
{\bf Varying Gas Fraction}
\\
\noindent
We also performed a test in which we varied the initial gas fraction of the
disk. In our fiducial run, we set the surface density of the gas disk from
$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{kpc}^2$.
We also ran with surface densities of $15$, $10$, and
$5\,\Msun/\textrm{kpc}^2$. These correspond to initial gas fractions of
approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
shown in Extended Data Fig.~\ref{fig:fgas} We find that the bar in disks with
initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{kpc}$ evolve
with a constant pattern speed while the bar in a disk with initial surface
density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.

As a result, we conclude that for the disk, bar, and halo properties
considered in this work, a gas fraction of only approximately $5\%$ is
necessary in order for the proposed stabilizing mechanism to operate. Furthermore, we co
